---
title: "Simulating data for the Ontario restaurant closures experiment"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2:
    toc: no
subtitle: "SUBTITLE"
abstract: "This document explains how data for the Ontario restaurant closures experiment was simulated, including the research that was used for parameterization and the methods used to randomize data."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in projects

```

```{r}
# read in the data
# NOTE: scripts 01_scrape_health_depts and 02_sampling_frame must have been run already

table_for_surveys <- read_csv(here::here("outputs/table_for_surveys.csv"))

survey_1 <- table_for_surveys %>%
  select(ID, group) %>%
  arrange(ID)

# count the numbers for treat, control, and total
num_treat <- first(count(data %>%
                           filter(group == "treatment")))

num_control <- first(count(data %>% 
                             filter(group == "control")))

num_total <- first(count(data))

survey_2 <- survey_1

survey_1
```



# Survey questions that require simulated data

# Type of service provided
LORENA
## Survey 1
```{r}
survey_1 <- survey_1 %>%
```

## Survey 2
```{r}
survey_2 <- survey_2 %>%
```


# Demographic traits of owners
AMY AND LALA

## Survey 1
```{r}
survey_1 <- survey_1 %>%
```

## Survey 2

Based on:https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3310028101
there seems to be no significant difference in business losses for disabled owners.

```{r}
survey_2 <- survey_2 %>%
  
```


# Shutdowns

https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3310027601
43.2% shut down temporarily

https://nypost.com/2020/06/30/more-than-half-of-us-restaurants-closed-due-to-covid-19-study/
3% closed permanently - but that was over far more than two weeks.

# Survey 2
```{r}
set.seed(24)
survey_2 <- bind_rows(
  survey_2 %>%
    filter(group == "treatment") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"), 
                                                            size = num_treat, 
                                                            prob = c(0.558, 0.432, 0.01),
                                                            replace = TRUE))
  ,
  survey_2 %>%
    filter(group == "control") %>%
    mutate(shutdown = sample(c("none", "temporary", "permanent"), 
                                                          size = num_control, 
                                                          prob = c(.99, 0.005, 0.005),
                                                          replace = TRUE))
) %>%
  arrange(ID)
  
```


# Revenue
From https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310022201&pickMembers%5B0%5D=2.1&pickMembers%5B1%5D=3.423
25,836 employers

From https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=3310022301&pickMembers%5B0%5D=2.423
6,968 non-employers

So 32,804 businesses in 2019

https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2110017101&pickMembers%5B0%5D=1.7&pickMembers%5B1%5D=2.2&cubeTimeFrame.startYear=2015&cubeTimeFrame.endYear=2019&referencePeriods=20150101%2C20190101
full-service restaurants have 13,456,600,000 in revenue in 2019.
https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2110017101&pickMembers%5B0%5D=1.7&pickMembers%5B1%5D=2.3&cubeTimeFrame.startYear=2015&cubeTimeFrame.endYear=2019&referencePeriods=20150101%2C20190101
limited-service eating places have 14,082,700,000 in revenue in 2019.

so averaging 839,510 in annual revenue.

69,959 average per month.

From http://www.mbel.io/2019/08/23/kaggle-restaurant-revenue-prediction/
we can see the distribution looks like an F distribution, or a log normal distribution.
rf(n, df1, df2, ncp) = rf(num_rest, 10, 5) should have a mean of 5/3
so we would multiply by 41,975 to get the mean where we want.

## Survey 1
```{r}
set.seed(24)
survey_1 <- survey_1 %>%
  mutate(revenue = 41975 * rf(num_total,10,5))
```

need to account for total and partial shutdowns: assume that total shutdowns cut revenue 75%, and partial cut revenue 50%

from https://www.eater.com/2020/3/24/21184301/restaurant-industry-data-impact-covid-19-coronavirus
revenues went from +5 to -35 when the shutdowns started. loss of -38. 
but takeout only places need -0
dine-in only places -100*14/31 = -45
and combo places need -38*14/31 = -17

####################### TODO: account for minority owned businesses

## Survey 2
```{r}
survey_2 <- survey_2 %>%
  mutate(revenue = 41975 * rf(num_total,10,5))

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(shutdown == "none") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(shutdown == "temporary") %>%
      mutate(revenue = 0.50 * revenue)
    ,
    survey_2 %>%
      filter(shutdown == "permanent") %>%
      mutate(revenue = 0.25 * revenue)
  ) %>%
  arrange(ID)

####################33 THIS WONT WORK UNTIL LORENA"S PART IS ADDED ON TYPE
survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(type == "dinein" & group == "treatment") %>%
      mutate(revenue = 0.55 * revenue)
    ,
    survey_2 %>%
      filter(type == "both" & group == "treatment") %>%
      mutate(revenue = 0.83 * revenue)
    ,
    survey_2 %>%
      filter(type == "takeout" & group == "treatment") %>%
      mutate(revenue = revenue)
    ,
    survey_2 %>%
      filter(group == "control")
  ) %>%
  arrange(ID)

survey_2 <-
  bind_rows(
    survey_2 %>%
      filter(minority == "yes" & group == "treatment") %>%
      mutate(revenue = 0.9 * revenue)
    ,
    survey_2 %>%
      filter(minority == "no" & group == "treatment")
    ,
    survey_2 %>%
      filter(group == "control")
  )

```




# Number of employees
LALA
## Survey 1
```{r}
survey_1 <- survey_1 %>%
```

## Survey 2
```{r}
survey_2 <- survey_2 %>%
```





# Applications for government assistance
LORENA
## Survey 1
```{r}
survey_1 <- survey_1 %>%
```

## Survey 2
```{r}
survey_2 <- survey_2 %>%
```




\newpage

# References