---
title: "Simulating data for the Ontario restaurant closures experiment"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2:
    toc: no
subtitle: "SUBTITLE"
abstract: "This document explains how data for the Ontario restaurant closures experiment was simulated, including the research that was used for parameterization and the methods used to randomize data."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in projects

```

```{r}
# read in the data
all_units_data <- read_csv(here::here("outputs/data/all_units_data.csv"))

# assign treatment and control based on health unit random selection
all_units_data <- all_units_data %>%
  mutate(group = case_when(unit == "brant" ~ "treatment",
                           unit == "peel" ~ "treatment",
                           unit == "sudbury" ~ "treatment",
                           unit == "hamilton" ~ "control",
                           unit == "southwestern" ~ "control",
                           unit == "chatham" ~ "control"
                           ))

all_treat <- all_units_data %>%
  filter(group == "treatment")

all_control <- all_units_data %>%
  filter(group == "control")

# count the numbers for treat, control, and total
num_treat <- first(count(all_treat))

num_control <- first(count(all_control))

num_total <- first(count(all_units_data))


# randomly select 10% of total to be surveyed for #1, and recount.
set.seed(24)
survey_1 <- bind_rows(
  all_treat %>%
    mutate(surveyed = sample(x = c("yes", "no"),
                             size = num_treat,
                             prob = c(0.1, 0.9),
                             replace = TRUE
                             )) %>%
    filter(surveyed == "yes"),
  all_control %>%
    mutate(surveyed = sample(x = c("yes", "no"),
                             size = num_control,
                             prob = c(0.1, 0.9),
                             replace = TRUE
                             )) %>%
    filter(surveyed == "yes")
)
  

num_treat_1 <- first(count(survey_1 %>%
                   filter(group == "treatment")))

num_control_1 <- first(count(survey_1 %>%
                   filter(group == "control")))

num_total_1 <- first(count(survey_1))


# randomly select 10% of total to be surveyed for #2, and recount.
set.seed(25)
survey_2 <- bind_rows(
  all_treat %>%
  mutate(surveyed = sample(x = c("yes", "no"),
                           size = num_treat,
                           prob = c(0.1, 0.9),
                           replace = TRUE
                           )) %>%
  filter(surveyed == "yes"),
  all_control %>%
  mutate(surveyed = sample(x = c("yes", "no"),
                           size = num_control,
                           prob = c(0.1, 0.9),
                           replace = TRUE
                           )) %>%
  filter(surveyed == "yes")
)

num_treat_2 <- first(count(survey_2 %>%
                   filter(group == "treatment")))

num_control_2 <- first(count(survey_2 %>%
                   filter(group == "control")))

num_total_2 <- first(count(survey_2))

```



# Survey questions that require simulated data

# Type of service provided
LORENA
## Survey 1
```{r}

```

## Survey 2
```{r}

```


# Number of employees
LALA
## Survey 1
```{r}

```

## Survey 2
```{r}

```


# Revenue
AMY

## Survey 1
```{r}

```

## Survey 2
```{r}

```


# Applications for government assistance
LORENA
## Survey 1
```{r}

```

## Survey 2
```{r}

```


# Demographic traits of owners
AMY AND LALA

## Survey 1
```{r}

```

## Survey 2

Based on:https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3310028101
there seems to be no significant difference in business losses for disabled owners.

```{r}

```

\newpage

# References